### 2017-10-26 Risk analysis/Text/Layout/Layout/chapter2_methods.tex
Start hydrocarbon traps text (methods)

        \section{Application in structural geological modeling}\label{sec:bayes_in_struc_modeling}
        In this work, these methods of Bayesian analysis and decision theory are applied in the field of geological structural modeling. The fundamental approach follows closely the research conducted by \citet{delaVarga2016} and builds upon their findings.\\
        According to them, structural geological modeling can be regarded as a forward problem and the elements of Bayesian inference can be specified in this context as follows:
        \begin{enumerate}
        	\item \textbf{Mathematical forward model ($M$)}: The connections between parameters $\theta$ and observed data $y$ are defined in such mathematical models. According to \citet{delaVarga2016} (and GREENSTONE PAPER, CITE?), the realization of a geological model $M$ can be regarded as a direct function of a set of input parameters:
        	\begin{equation}
        	M = f(\vec{x}, \phi_i, k_j, \alpha_k, \beta_l)\label{eq:forward_model},
        	\end{equation}
        	where $\phi_i$ is a function of position $\vec{x}$, more precisely an interpolation function, to which respective additional parameters are given by $\alpha_k$. Available primary geological information, such as positions of layer interfaces and dips in the subsurface, is represented by $k_j$. A topological description, such as the relationships between faults and layer surfaces, is given by $\beta_l$. An essential aspect of this function is that it allows for a full automation of the modeling step, so that the consequences of a change in an input parameter are realized directly without the need for any further manual inspection or interaction (cite GREENSTONE paper). The modeling step and respective interpolation used in this work is presented in Section \ref{sec:struc_geo_modeling} below.
        	\item \textbf{Model parameters ($\theta$)}: These model-defining parameters can be deterministic or stochastic. In the latter case, they are uncertain parameters to which a probability distribution is assigned. Regarding Equation \ref{eq:forward_model} above, these can be any of the factors $\vec{x}$, $k_j$, $\alpha_k$ and $\beta_l$, which define the realization of the forward model over $\phi_i$.
        	\item \textbf{Observed data ($y$)}: This is any type of additional information that can be related to the forward modeling results and might possibly be used to reduce uncertainty. Such data can be gained by measurements and observations, for example by core sampling, well-log analysis or seismic acquisition. 
        	\item \textbf{Likelihood functions ($p(y|\theta)$}: Links between the previous parameters $\theta$ and the additional data $y$ are established by these functions in way that they reflect the likelihood of the parameter states given the observations. They mathematically defined in the same way as probability functions, but are a function of the data $y$, instead of the parameters $\theta$ \citep{delaVarga2016}.
        \end{enumerate}

@@

        Thereby, the algorithm assigns high probabilities to high-density points and low probabilities to low-density points, so that the chain state is moved accordingly \citep{delaVarga2016}.\\
        Metropolis methods are furthermore defined by the step size scale factor that is chosen. While large steps are good for exploration of the space and mixture in the chain, acceptance rates are low. Small steps have better acceptance rates, but lead to slower exploration and convergence of the algorithm \citep{delaVarga2016}.\\
        For the iterative sampling, Adaptive Metropolis (AM) by \citet{haario2001adaptive} is used in this work. It adapts the traditional Metropolis-Hastings by incorporating the ability of continuous step size tuning during convergence, by taking into account the full information saved along the process. This is achieved by generating a covariance matrix that is updated every iteration. The adaptive nature of the process enables fast convergence for non-linear distributions while maintaining ergodicity \citet{haario2001adaptive, delaVarga2016}. Its suitability for multi-dimensional distribution spaces make it an excellent method for dealing with complex models such as the structural geological models in this work \citep{schaaf2017}.\\
        A large enough number of iterations $T$ has to be chosen, so that a reliable and statistically significant exploration of the parameter space is assured. This is primarily dependent on the rate of convergence towards the true distribution. Considering empirical Bayes methods, in which either priors or likelihoods stem from empirical data, as is assumed for the models in this work, convergence can be expected to be reached almost immediately. However, for high dimensional problems more iterations are required, in order to ensure an accurate representation of posterior distributions (cite Miguel GREENSTONE!). Different realizations of the full 3D geological model are then constructed on the basis of the resulting posterior distributions, using an implicit modeling step described in the following.

@@

        \subsection{Economic significance and reservoir valuation}\label{sec:Reservoir_values}
        In this work, the generated structural models are to be assigned an economic significance in order to approach a real setting, in which hypothetical actors could have an interest in taking a decision, and thus, to evaluate the influence of the Bayesian inference of the decision-making process. Changes in Bayes actions are considered as a measure for this effect.\\
        For this purpose, it is assumed that a hydrocarbon reservoir system is represented by each geological model realization. These are typically comprised stratigraphically of at least one potential reservoir formation, succeeded by at least one sealing layer above. Additionally, a trapping mechanism is normally required (source?).\\
        The resulting reservoirs are to be valuated in a way that respective estimations on the basis of loss functions can be conducted, i.e. Bayesian decision theory is applicable. In the typical petroleum industrial setting, the main question is one of how much of the resource can be produced and how high the return on investment will be \citep{dean2007volumetric}.

>>>>

        \section{Application in structural geological modeling}\label{sec:bayes_in_struc_modeling}
        In this work, these methods of Bayesian analysis and decision theory are applied in the field of geological structural modeling. The fundamental approach follows closely the research conducted by \citet{delaVarga2016} and builds upon their findings.\\
        According to them, structural geological modeling can be regarded as a forward problem and the elements of Bayesian inference can be specified in this context as follows:
        \begin{enumerate}
        	\item \textbf{Mathematical forward model ($M$)}: The connections between parameters $\theta$ and observed data $y$ are defined in such mathematical models. According to \citet{delaVarga2016}, the realization of a geological model $M$ can be regarded as a direct function of a set of input parameters:
        	\begin{equation}
        	M = f(\vec{x}, \phi_i, k_j, \alpha_k, \beta_l)\label{eq:forward_model},
        	\end{equation}
        	where $\phi_i$ is a function of position $\vec{x}$, more precisely an interpolation function, to which respective additional parameters are given by $\alpha_k$. Available primary geological information, such as positions of layer interfaces and dips in the subsurface, is represented by $k_j$. A topological description, such as the relationships between faults and layer surfaces, is given by $\beta_l$. An essential aspect of this function is that it allows for a full automation of the modeling step, so that the consequences of a change in an input parameter are realized directly without the need for any further manual inspection or interaction \citep{wellmann2017sandstone}. The modeling step and respective interpolation used in this work is presented in Section \ref{sec:struc_geo_modeling} below.
        	\item \textbf{Model parameters ($\theta$)}: These model-defining parameters can be deterministic or stochastic. In the latter case, they are uncertain parameters to which a probability distribution is assigned. Regarding Equation \ref{eq:forward_model} above, these can be any of the factors $\vec{x}$, $k_j$, $\alpha_k$ and $\beta_l$, which define the realization of the forward model over $\phi_i$.
        	\item \textbf{Observed data ($y$)}: This is any type of additional information that can be related to the forward modeling results and might possibly be used to reduce uncertainty. Such data can be gained by measurements and observations, for example by core sampling, well-log analysis or seismic acquisition. 
        	\item \textbf{Likelihood functions ($p(y|\theta)$}: Links between the previous parameters $\theta$ and the additional data $y$ are established by these functions in way that they reflect the likelihood of the parameter states given the observations. They mathematically defined in the same way as probability functions, but are a function of the data $y$, instead of the parameters $\theta$ \citep{delaVarga2016}.
        \end{enumerate}

@@

        Thereby, the algorithm assigns high probabilities to high-density points and low probabilities to low-density points, so that the chain state is moved accordingly \citep{delaVarga2016}.\\
        Metropolis methods are furthermore defined by the step size scale factor that is chosen. While large steps are good for exploration of the space and mixture in the chain, acceptance rates are low. Small steps have better acceptance rates, but lead to slower exploration and convergence of the algorithm \citep{delaVarga2016}.\\
        For the iterative sampling, Adaptive Metropolis (AM) by \citet{haario2001adaptive} is used in this work. It adapts the traditional Metropolis-Hastings by incorporating the ability of continuous step size tuning during convergence, by taking into account the full information saved along the process. This is achieved by generating a covariance matrix that is updated every iteration. The adaptive nature of the process enables fast convergence for non-linear distributions while maintaining ergodicity \citet{haario2001adaptive, delaVarga2016}. Its suitability for multi-dimensional distribution spaces make it an excellent method for dealing with complex models such as the structural geological models in this work \citep{schaaf2017}.\\
        A large enough number of iterations $T$ has to be chosen, so that a reliable and statistically significant exploration of the parameter space is assured. This is primarily dependent on the rate of convergence towards the true distribution. Considering empirical Bayes methods, in which either priors or likelihoods stem from empirical data, as is assumed for the models in this work, convergence can be expected to be reached almost immediately. However, for high dimensional problems more iterations are required, in order to ensure an accurate representation of posterior distributions \citep{wellmann2017sandstone}. Different realizations of the full 3D geological model are then constructed on the basis of the resulting posterior distributions, using an implicit modeling step described in the following.

@@

        \subsection{Economic significance and reservoir valuation}\label{sec:Reservoir_values}
        In this work, the generated structural models are to be assigned an economic significance in order to approach a real setting, in which hypothetical actors could have an interest in taking a decision, and thus, to evaluate the influence of the Bayesian inference of the decision-making process. Changes in Bayes actions are considered as a measure for this effect.      For this purpose, it is assumed that a hydrocarbon reservoir system is represented by each geological model realization.

        \subsubsection{Reservoirs and structural hydrocarbon traps}
        Such hydrocarbon are typically comprised of:
        \begin{enumerate}
        	\item
        \end{enumerate} \citep{dolson2016basics}
        These are typically comprised stratigraphically of at least one potential reservoir formation, succeeded by at least one sealing layer above. Additionally, a trapping mechanism is normally required (source?).\\
        The resulting reservoirs are to be valuated in a way that respective estimations on the basis of loss functions can be conducted, i.e. Bayesian decision theory is applicable. In the typical petroleum industrial setting, the main question is one of how much of the resource can be produced and how high the return on investment will be \citep{dean2007volumetric}.
